---
title: "AnalysisCOVID19Data"
author: "Richard Heal"
date: "8/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#######################################################################
# Load in the data
#######################################################################
DataFile <- "COVIDAquaculture_Data_RawDATA.csv"
RawData <- read.csv(DataFile, header = TRUE, skip = 2, stringsAsFactors = TRUE)

##########################################
# Load in the consumer data
##########################################
ConsumerDataFile <- "COVIDAquaculture_ConsumerData.csv"
ConsumerData <- read.csv(ConsumerDataFile, header = TRUE, skip = 2, stringsAsFactors = TRUE)

```

## Analysis of Aquaculture in Bangladesh Pre and Post COVID 19

This document details the analysis of farm survey data taken pre and post COVID 19.

```{r Effect on Labour BASIC STATS, include=FALSE}

#############################################################
#
# Analysis on the effect of COVID on labour for the farms
#
# Here the number of workers and the cost of labour for 
# permanent, seasonal and non-technical staff
#
# 
#############################################################
# Overall figures
#############################################################
library(data.table)
RawData <- setDT(RawData)
RawData[, TotLab_1 := (PermLab_1 + SeasLab_1 + NonAqLab_1)]
RawData[, TotLab_2 := (PermLab_2 + SeasLab_2 + NonAqLab_2)]
RawData[, DiffLab := (TotLab_2 - TotLab_1)]
RawData[, DiffPerm := (PermLab_2 - PermLab_1)]
RawData[, DiffSeas := (SeasLab_2 - SeasLab_1)]
RawData[, DiffNonAq := (NonAqLab_2 - NonAqLab_1)]
Farms_MoreStaff <- sum(RawData$DiffLab > 0)
Farms_LessStaff <- sum(RawData$DiffLab < 0)
Farms_SameStaff <- sum(RawData$DiffLab == 0)

Mean_StaffChange <- mean(RawData$DiffLab, na.rm = TRUE)
Median_StaffChange <- median(RawData$DiffLab, na.rm = TRUE)

Total_Perm_1 <- sum(RawData$PermLab_1)
Total_Seas_1 <- sum(RawData$SeasLab_1)
Total_NonAq_1 <- sum(RawData$NonAqLab_1)
Total_Perm_2 <- sum(RawData$PermLab_2)
Total_Seas_2 <- sum(RawData$SeasLab_2)
Total_NonAq_2 <- sum(RawData$NonAqLab_2)


```
  
### Effect on labour - employee numbers
  
The overall effect on the labour numbers:

  Farms with more staff: **`r Farms_MoreStaff`**  
  Farms with less staff: **`r Farms_LessStaff`**  
  Farms with the same number of staff employed: **`r Farms_SameStaff`**  

So most farms reported a decrease in employee numbers.  The mean change in employee number was **`r Mean_StaffChange`** and the median change was **`r Median_StaffChange`**.

Before COVID19 the farms employed **`r Total_Perm_1`** permanent staff, **`r Total_Seas_1`** seasonal staff and **`r Total_NonAq_1`** non-technical staff.

After COVID19 the farms employed **`r Total_Perm_2`** permanent staff, **`r Total_Seas_2`** seasonal staff and **`r Total_NonAq_2`** non-technical staff.  
  
  Difference in permanent staff numbers: **`r Total_Perm_2 - Total_Perm_1`**  
  Difference in seasonal staff numbers: **`r Total_Seas_2 - Total_Seas_1`**  
  Difference in non-technical staff numbers: **`r Total_NonAq_2 - Total_NonAq_1`**  
  
The effect was appears to be a reduction in permanent and non-technical staff numbers with seasonal numbers remaining consistent. However, it is a little more complicated than that.  

For example, of all the farms **`r sum(RawData$DiffPerm < 0)`** reduced permanent staff, **`r sum(RawData$DiffPerm > 0)`** increased permanent staff and **`r sum(RawData$DiffPerm == 0)`** kept the same number of permanent staff. 

INPUT THE RANGE OF VALUES.

Those farms that reduced permanent staff did so by a mean of **`r mean(RawData$DiffPerm[which(RawData$DiffPerm < 0)])`** employees with a median of **`r median(RawData$DiffPerm[which(RawData$DiffPerm < 0)])`** and a range of **`r min(RawData$DiffPerm[which(RawData$DiffPerm < 0)])`** to **`r max(RawData$DiffPerm[which(RawData$DiffPerm < 0)])`**. Those farms that increased permanent staff did so by a mean of **`r mean(RawData$DiffPerm[which(RawData$DiffPerm > 0)])`** employees with a median of **`r median(RawData$DiffPerm[which(RawData$DiffPerm > 0)])`**.

For seasonal staff, of all the farms **`r sum(RawData$DiffSeas < 0)`** reduced seasonal staff, **`r sum(RawData$DiffSeas > 0)`** increased seasonal staff and **`r sum(RawData$DiffSeas == 0)`** kept the same number of seasonal staff.

Those farms that reduced seasonal staff did so by a mean of **`r mean(RawData$DiffSeas[which(RawData$DiffSeas < 0)])`** employees with a median of **`r median(RawData$DiffSeas[which(RawData$DiffSeas < 0)])`** and a range of **`r min(RawData$DiffSeas[which(RawData$DiffSeas < 0)])`** to **`r max(RawData$DiffSeas[which(RawData$DiffSeas < 0)])`**. Those farms that increased seasonal staff did so by a mean of **`r mean(RawData$DiffSeas[which(RawData$DiffSeas > 0)])`** employees with a median of **`r median(RawData$DiffSeas[which(RawData$DiffSeas > 0)])`**.

For non-technical staff, of all the farms **`r sum(RawData$DiffNonAq < 0)`** reduced non-technical staff, **`r sum(RawData$DiffNonAq > 0)`** increased non-technical staff and **`r sum(RawData$DiffNonAq == 0)`** kept the same number of non-technical staff.

Those farms that reduced non-technical staff did so by a mean of **`r mean(RawData$DiffNonAq[which(RawData$DiffNonAq < 0)])`** employees with a median of **`r median(RawData$DiffNonAq[which(RawData$DiffNonAq < 0)])`** and a range of **`r min(RawData$DiffNonAq[which(RawData$DiffNonAq < 0)])`** to **`r max(RawData$DiffNonAq[which(RawData$DiffNonAq < 0)])`**. Those farms that increased non-technical staff did so by a mean of **`r mean(RawData$DiffNonAq[which(RawData$DiffNonAq > 0)])`** employees with a median of **`r median(RawData$DiffNonAq[which(RawData$DiffNonAq > 0)])`**.

The trend in seasonal staff was either a reduction or increase by a single employee, or no change occurring. Of those **`r sum(RawData$DiffSeas > 0)`** farms increasing the seasonal staff, **`r sum(RawData$DiffPerm[which(RawData$DiffSeas > 0)] < 0)`** farms had decreased their permanent staff, **`r sum(RawData$DiffNonAq[which(RawData$DiffSeas > 0)] < 0)`** farms had decreased their non-technical staff, and **`r sum(RawData$DiffNonAq[which(RawData$DiffSeas > 0)]< 0 & RawData$DiffPerm[which(RawData$DiffSeas > 0)] < 0)`** reduced both permanent and non-technical staff. **`r sum(RawData$DiffPerm[which(RawData$DiffSeas > 0)] == 0)`** farms kept the same number of permanent staff, and  **`r sum(RawData$DiffNonAq[which(RawData$DiffSeas > 0)] == 0)`** kept the same number of non-technical staff, and **`r sum(RawData$DiffNonAq[which(RawData$DiffSeas > 0)] == 0 & RawData$DiffPerm[which(RawData$DiffSeas > 0)] == 0)`** kept the same number of permanent **and** non-technical staff, whilst increasing seasonal employees.  

```{r Effect on Labour COSTS BASIC STATS, include=FALSE}
#########################################################
#
# Here some basic figures on the cost of labour changes
#
#########################################################
MeanPermCost_1 <- mean(RawData$PermLabCOST_1, na.rm = TRUE)
MeanPermCost_2 <- mean(RawData$PermLabCOST_2, na.rm = TRUE)
MeanSeasCost_1 <- mean(RawData$SeasLabCOST_1, na.rm = TRUE)
MeanSeasCost_2 <- mean(RawData$SeasLabCOST_2, na.rm = TRUE)
MeanNonAqCost_1 <- mean(RawData$NonAqLabCOST_1, na.rm = TRUE)
MeanNonAqCost_2 <- mean(RawData$NonAqLabCOST_2, na.rm = TRUE)

Mean_Total_1 <- (MeanPermCost_1 + MeanSeasCost_1 + MeanNonAqCost_1)/3
Mean_Total_2 <- (MeanPermCost_2 + MeanSeasCost_2 + MeanNonAqCost_2)/3

############################################################
# Paired t-tests
############################################################
Perm_Cost_TTest <- t.test(RawData$PermLabCOST_2,
                          RawData$PermLabCOST_1,
                          alternative = "less",
                          paired = TRUE,
                          na.action = na.omit)

Seas_Cost_TTest <- t.test(RawData$SeasLabCOST_2,
                          RawData$SeasLabCOST_1,
                          alternative = "less",
                          paired = TRUE,
                          na.action = na.omit)

NonAq_Cost_TTest <- t.test(RawData$NonAqLabCOST_2,
                          RawData$NonAqLabCOST_1,
                          alternative = "less",
                          paired = TRUE,
                          na.action = na.omit)
```

### Effect on labour - costs of employment per day
  
The overall effect on the labour costs:

The mean labour cost pre-COVID was **`r Mean_Total_1`** and was **`r Mean_Total_2`** post-COVID suggesting a small decrease in labour cost due to COVID. For each of the employment types the means were:  
  
  Pre-covid Permanent Labour cost mean: **`r MeanPermCost_1`**  
  Post-covid Permanent Labour cost mean: **`r MeanPermCost_2`**  
  
  Pre-covid Seasonal Labour cost mean: **`r MeanSeasCost_1`**  
  Post-covid Seasonal Labour cost mean: **`r MeanSeasCost_2`**  
  
  Pre-covid Non-technical Labour cost mean: **`r MeanNonAqCost_1`**  
  Post-covid Non-technical Labour cost mean: **`r MeanNonAqCost_2`**  

Therefore, there has been reduction in all employment sector costs due to COVID. These values are the costs per employee and therefore this suggests a decrease in the wages for sector employees.

Assuming the costs are normally distributed (which is a fair assumption to make) we can check whether there is a statistical difference between the costs per employee. This will be a pairwise test using farms with before and after data.

#### T-test for Labour Costs

To test whether there was a significant difference in the costs per employee pre and post COVID 19 a one-tailed paired t-test was performed on the data.

For permanent employee costs:  
        P-value: **`r Perm_Cost_TTest$p.value`**  
        Test Statistic: **`r Perm_Cost_TTest$statistic`**  
        Degrees of Freedom: **`r Perm_Cost_TTest$parameter`**  
        Estimated difference in means: **`r Perm_Cost_TTest$estimate`**  
          
For seasonal employee costs:    
        P-value: **`r Seas_Cost_TTest$p.value`**  
        Test Statistic: **`r Seas_Cost_TTest$statistic`**  
        Degrees of Freedom: **`r Seas_Cost_TTest$parameter`**  
        Estimated difference in means: **`r Seas_Cost_TTest$estimate`**  
          
For non-technical employee costs:    
        P-value: **`r NonAq_Cost_TTest$p.value`**  
        Test Statistic: **`r NonAq_Cost_TTest$statistic`**  
        Degrees of Freedom: **`r NonAq_Cost_TTest$parameter`**  
        Estimated difference in means: **`r NonAq_Cost_TTest$estimate`**  

The results suggest a significant decrease in the costs of labour for seasonal and non-technical employees with a decrease of 0.5USD per day (**`r Seas_Cost_TTest$estimate`** for seasonal staff and **`r NonAq_Cost_TTest$estimate`** for non-technical staff). There was no significant difference in the costs for permanent staff.  

```{r Effect on Labour SCATTER PLOTS, include=FALSE}
##############################################################
#Going to produce scatter plots pre and post covid
###############################################################
library(plotly)
Offset = 0.1
MaxX = 6 + Offset
MaxY = 6+ Offset
MinX = -Offset
MinY = -Offset
SymbolSize = 7

# Set axis titles
x1 <- list(title = list(text = "Employee Number Pre-COVID19"), 
                      range = c(MinX,MaxX))
y1 <- list(title = list(text = "Employee Number Post-COVID19"),
                      range = c(MinY,MaxY))

PermWorkersPerFarmPlot <- plot_ly(data = RawData, type = 'scatter', mode = 'markers+line') %>%
  add_trace(name = 'NoChangeLine',
            mode = 'line',
            fill = 'tozeroy', fillcolor = 'rgba(240,240,240, 0.8)',
            showlegend = FALSE,
            line = list(color = 'rgb(200,200,200)', width = 0.5, dash = '5px'),
            x = c(0,MaxX),
            y = c(0,MaxY)) %>%
  add_trace(mode = 'markers',
            marker = list(color = 'rgb(0,0,0)', size = SymbolSize, symbol = 'triangle-up'),
            name = 'Permanent Labour Numbers',
            x = ~jitter(PermLab_1, amount = 0),
            y = ~jitter(PermLab_2, amount = 0)) %>%
  layout(showlegend = FALSE,
         xaxis = list(title = list(text = "Permanent Employee Number Pre-COVID19"), 
                      range = c(MinX,MaxX)),
         yaxis = y1)
SeasWorkersPerFarmPlot <- plot_ly(data = RawData, type = 'scatter', mode = 'markers+line') %>%
  add_trace(name = 'NoChangeLine',
            mode = 'line',
            fill = 'tozeroy', fillcolor = 'rgba(240,240,240, 0.8)',
            showlegend = FALSE,
            line = list(color = 'rgb(200,200,200)', width = 0.5, dash = '5px'),
            x = c(0,MaxX),
            y = c(0,MaxY)) %>%
  add_trace(mode = 'markers',
            marker = list(color = 'rgb(0,0,0)', size = SymbolSize, symbol = 'triangle-down'),
            name = 'Seasonal Labour',
            x = ~jitter(SeasLab_1, amount = 0),
            y = ~jitter(SeasLab_2, amount = 0)) %>%
  layout(showlegend = FALSE,
         xaxis = list(title = list(text = "Seasonal Employee Number Pre-COVID19"), 
                      range = c(MinX,MaxX)),
         yaxis = y1)
NonAqWorkersPerFarmPlot <- plot_ly(data = RawData, type = 'scatter', mode = 'markers+line') %>%
  add_trace(name = 'NoChangeLine',
            mode = 'line',
            fill = 'tozeroy', fillcolor = 'rgba(240,240,240, 0.8)',
            showlegend = FALSE,
            line = list(color = 'rgb(200,200,200)', width = 0.5, dash = '5px'),
            x = c(0,MaxX),
            y = c(0,MaxY)) %>%
  add_trace(mode = 'markers',
            marker = list(color = 'rgb(0,0,0)', size = SymbolSize, symbol = 'circle'),
            name = 'Non-specialist Labour',
            x = ~jitter(NonAqLab_1, amount = 0),
            y = ~jitter(NonAqLab_2, amount = 0)) %>%
  layout(showlegend = FALSE,
         xaxis = list(title = list(text = "Non-technical Employee Number Pre-COVID19"), 
                      range = c(MinX,MaxX)),
         yaxis = y1)
WorkersPerFarmPlot <- subplot(PermWorkersPerFarmPlot,
                              SeasWorkersPerFarmPlot,
                              NonAqWorkersPerFarmPlot,
                              nrows = 1,
                              shareY = TRUE,
                              shareX = TRUE,
                              titleX = TRUE,
                              titleY = TRUE)

####################################################################################
#
# Repeat for costs
#
#####################################################################################
Offset = 0.1
MaxX = 6 + Offset
MaxY = 6+ Offset
MinX = 2 - Offset
MinY = 2 - Offset
SymbolSize = 7

# Set axis titles
x1 <- list(title = list(text = "Labour Cost Pre-COVID19 (USD)"), 
                      range = c(MinX,MaxX))
y1 <- list(title = list(text = "Labour Cost Post-COVID19"),
                      range = c(MinY,MaxY),
           showline = T)

PermWorkersCOSTPerFarmPlot <- plot_ly(data = RawData, type = 'scatter', mode = 'markers+line') %>%
  add_trace(name = 'NoChangeLine',
            mode = 'line',
            fill = 'tozeroy', fillcolor = 'rgba(240,240,240, 0.8)',
            showlegend = FALSE,
            line = list(color = 'rgb(200,200,200)', width = 0.5, dash = '5px'),
            x = c(0,MaxX),
            y = c(0,MaxY)) %>%
  add_trace(mode = 'markers',
            marker = list(color = 'rgb(0,0,0)', size = SymbolSize, symbol = 'triangle-up-open'),
            name = 'Permanent Labour Numbers',
            x = ~jitter(PermLabCOST_1, amount = 0),
            y = ~jitter(PermLabCOST_2, amount = 0)) %>%
  layout(showlegend = FALSE,
         yaxis = y1,
         xaxis = list(title = list(text = "Permanent Labour Cost Pre-COVID19 (USD)"),
                      range = c(MinY,MaxY),
                      showline = T))
SeasWorkersCOSTPerFarmPlot <- plot_ly(data = RawData, type = 'scatter', mode = 'markers+line') %>%
  add_trace(name = 'NoChangeLine',
            mode = 'line',
            fill = 'tozeroy', fillcolor = 'rgba(240,240,240, 0.8)',
            showlegend = FALSE,
            line = list(color = 'rgb(200,200,200)', width = 0.5, dash = '5px'),
            x = c(0,MaxX),
            y = c(0,MaxY)) %>%
  add_trace(mode = 'markers',
            marker = list(color = 'rgb(0,0,0)', size = SymbolSize, symbol = 'triangle-down-open'),
            name = 'Seasonal Labour',
            x = ~jitter(SeasLabCOST_1, amount = 0),
            y = ~jitter(SeasLabCOST_2, amount = 0)) %>%
  layout(showlegend = FALSE,
         yaxis = y1,
         xaxis = list(title = list(text = "Seasonal Labour Cost Pre-COVID19 (USD)"),
                      range = c(MinY,MaxY),
                      showline = T))
NonAqWorkersCOSTPerFarmPlot <- plot_ly(data = RawData, type = 'scatter', mode = 'markers+line') %>%
  add_trace(name = 'NoChangeLine',
            mode = 'line',
            fill = 'tozeroy', fillcolor = 'rgba(240,240,240, 0.8)',
            showlegend = FALSE,
            line = list(color = 'rgb(200,200,200)', width = 0.5, dash = '5px'),
            x = c(0,MaxX),
            y = c(0,MaxY)) %>%
  add_trace(mode = 'markers',
            marker = list(color = 'rgb(0,0,0)', size = SymbolSize, symbol = 'circle-open'),
            name = 'Non-specialist Labour',
            x = ~jitter(NonAqLabCOST_1, amount = 0),
            y = ~jitter(NonAqLabCOST_2, amount = 0)) %>%
  layout(showlegend = FALSE,
         yaxis = y1,
         xaxis = list(title = list(text = "Non-technical Labour Cost Pre-COVID19 (USD)"),
                      range = c(MinY,MaxY),
                      showline = T))
WorkersCOSTPerFarmPlot <- subplot(PermWorkersCOSTPerFarmPlot,
                              SeasWorkersCOSTPerFarmPlot,
                              NonAqWorkersCOSTPerFarmPlot,
                              nrows = 1,
                              shareY = TRUE,
                              shareX = TRUE,
                              titleX = TRUE,
                              titleY = TRUE)

WorkSubplot <- subplot(WorkersPerFarmPlot,
                       WorkersCOSTPerFarmPlot,
                       nrows = 2,
                       margin = 0.1,
                       titleX = TRUE,
                       titleY = TRUE)
```

### Scatter plots to show the effect

Here the effect of COVID19 is explored on the labour used to run the farms.

`r WorkersPerFarmPlot`

`r WorkersCOSTPerFarmPlot`

And, combined into a subplot

`r WorkSubplot`

### Summary

These results suggest:

 1. There are fewer permanent employees post-COVID 19 working on farms but the cost of their labour remains the same
 2. There are similar numbers of seasonal workers on farms post-COVID 19 and the cost of their labour is significantly reduced
 3. There are fewer non-technical staff on farms and their cost of labour is significantly less
 4. There is evidence to suggest some farms have reduced permanent and non-technical staff and replaced them with a single seasonal worker
 
```{r Total Number of Employees and Total Cost, include=FALSE}
#####################################################################################
# Looking at the effect on the total number of employees and the overall cost
#####################################################################################
# Insert the total cost for pre and post covid
RawData[, TotalCOST_Perm_1 := (PermLab_1 * PermLabCOST_1)]
RawData[, TotalCOST_Seas_1 := (SeasLab_1 * SeasLabCOST_1)]
RawData[, TotalCOST_NonAq_1 := (NonAqLab_1 * NonAqLabCOST_1)]
RawData[, TotalCOST_Perm_2 := (PermLab_2 * PermLabCOST_2)]
RawData[, TotalCOST_Seas_2 := (SeasLab_2 * SeasLabCOST_2)]
RawData[, TotalCOST_NonAq_2 := (NonAqLab_2 * NonAqLabCOST_2)]
RawData[, TotalCOST_1 := rowSums(.SD, na.rm = TRUE), .SDcols = c("TotalCOST_Perm_1", "TotalCOST_Seas_1", "TotalCOST_NonAq_1")]
RawData[, TotalCOST_2 := rowSums(.SD, na.rm = TRUE), .SDcols = c("TotalCOST_Perm_2", "TotalCOST_Seas_2", "TotalCOST_NonAq_2")]
RawData[, DiffTotalCOST := TotalCOST_2 - TotalCOST_1]


# Plot the difference in number versus the difference in cost

# Size of the bubble is the pre-covid19 total cost of labour and the colourscale is the post-covd19 cost
QuadrantTotalLabPlot <- plot_ly(data = RawData, type = 'scatter', mode = 'markers') %>%
  add_trace(name = 'Differences',
            mode = 'markers',
            x = ~DiffLab,
            y = ~DiffTotalCOST,
            marker = list(size = ~TotalCOST_1, color = ~TotalCOST_2, colorscale = 'Greys',
                          line = list(color = 'rgb(80,80,80)'),
                          sizemode = 'diameter', opacity = 0.5)) %>%
  layout(
    xaxis = list(title = "Difference in Staff Number",
                 range = c(-6,2)),
    yaxis = list(title = "Difference in total cost (USD)",
                 range = c(-35,5))
  )

# Try dumbbell for pre-covid total cost and bubble for post-covid cost and positioned on axis with total number of staff as size of bubble
PermColor <- "rgb(102,194,165)"
SeasColor <- "rgb(252,141,98)"
NonAqColor <- "rgb(141,160,203)"
MarkerSize <- 7

SingleTotalLabDumbbellPlot <- plot_ly(data = RawData, type = 'scatter', mode = 'markers') %>%
  add_segments(x = ~PermLabCOST_2, xend = ~PermLabCOST_1, y = ~(((as.numeric(X)-1) * 3) + 1), yend = ~(((as.numeric(X)-1) * 3) + 1), 
               showlegend = TRUE,
               name = 'Permanent',
               line = list(color = PermColor)) %>%
  add_markers(x = ~PermLabCOST_1,
              y = ~(((as.numeric(X)-1) * 3) + 1),
              name = "Pre-Covid",
              showlegend = TRUE,
              marker = list(size = MarkerSize, color = 'rgb(80,80,80)',
                          line = list(color = 'rgb(80,80,80)'),
                          sizemode = 'diameter', opacity = 0.5)) %>%
  add_markers(name = 'Post-Covid',
            showlegend = TRUE,
              x = ~PermLabCOST_2,
            y = ~(((as.numeric(X)-1) * 3) + 1),
            marker = list(size = MarkerSize, color = 'rgb(0,0,0)',
                          line = list(color = 'rgb(0,0,0)'),
                          sizemode = 'diameter', opacity = 0.5)) %>%
  add_segments(x = ~SeasLabCOST_2, xend = ~SeasLabCOST_1, y = ~(((as.numeric(X)-1) * 3) + 2), yend = ~(((as.numeric(X)-1) * 3) + 2), 
               showlegend = TRUE,
               name = 'Seasonal',
               line = list(color = SeasColor)) %>%
  add_markers(x = ~SeasLabCOST_1,
              y = ~(((as.numeric(X)-1) * 3) + 2),
              name = "Pre-Covid",
              showlegend = FALSE,
              marker = list(size = MarkerSize, color = 'rgb(80,80,80)',
                          line = list(color = 'rgb(80,80,80)'),
                          sizemode = 'diameter', opacity = 0.5)) %>%
  add_markers(name = 'Post-Covid',
              showlegend = FALSE,
            x = ~SeasLabCOST_2,
            y = ~(((as.numeric(X)-1) * 3) + 2),
            marker = list(size = MarkerSize, color = 'rgb(0,0,0)',
                          line = list(color = 'rgb(0,0,0)'),
                          sizemode = 'diameter', opacity = 0.5)) %>%
  add_segments(x = ~NonAqLabCOST_2, xend = ~NonAqLabCOST_1, y = ~(((as.numeric(X)-1) * 3) + 3), yend = ~(((as.numeric(X)-1) * 3) + 3), 
               showlegend = TRUE,
               name = 'Non-technical',
               line = list(color = NonAqColor)) %>%
  add_markers(x = ~NonAqLabCOST_1,
              y = ~(((as.numeric(X)-1) * 3) + 3),
              name = "Pre-Covid",
              showlegend = FALSE,
              marker = list(size = MarkerSize, color = 'rgb(80,80,80)',
                          line = list(color = 'rgb(80,80,80)'),
                          sizemode = 'diameter', opacity = 0.5)) %>%
  add_markers(name = 'Post-Covid',
              showlegend = FALSE,
            x = ~NonAqLabCOST_2,
            y = ~(((as.numeric(X)-1) * 3) + 3),
            marker = list(size = MarkerSize, color = 'rgb(0,0,0)',
                          line = list(color = 'rgb(80,80,80)'),
                          sizemode = 'diameter', opacity = 0.5)) %>%
  layout(
    xaxis = list(title = "Unit Labour Cost (USD)",
                 range = c(1,6), showline = TRUE),
    yaxis = list(title = "Farm",
                 range = c(0,122))
  )

############################################################
# Single plots to create a subplot
###########################################################
PreMarker = 'square-open'
PostMarker = 'circle'
PreMarkerColor = 'rgb(0,0,0)'
PostMarkerColor = 'rgb(0,0,0)'
SegmentLineColor = 'rgb(50,50,50)'

SingleTotalLabDumbbellPlot_PERM <- plot_ly(data = RawData, type = 'scatter', mode = 'markers') %>%
  add_segments(x = ~PermLabCOST_2, xend = ~PermLabCOST_1, y = ~as.numeric(X), yend = ~as.numeric(X), 
               showlegend = FALSE,
               name = 'Permanent',
               line = list(color = SegmentLineColor)) %>%
  add_markers(x = ~PermLabCOST_1,
              y = ~as.numeric(X),
              name = "Pre-Covid",
              showlegend = TRUE,
              marker = list(symbol = PreMarker,
                            size = MarkerSize, 
                          color = PreMarkerColor,
                          line = list(color = 'rgb(0,0,0)'),
                          sizemode = 'diameter', opacity = 0.5)) %>%
  add_markers(name = 'Post-Covid',
            showlegend = TRUE,
              x = ~PermLabCOST_2,
            y = ~as.numeric(X),
            marker = list(symbol = PostMarker,
                          size = MarkerSize, 
                          color = PostMarkerColor,
                          line = list(color = 'rgb(0,0,0)'),
                          sizemode = 'diameter', opacity = 0.5)) %>%
  layout(
    shapes = list(
      list(type = 'line', layer = 'bottom', 
                  y0 = 0, y1 = 41,
                  x0 = MeanPermCost_1, x1 = MeanPermCost_1,
                  line = list(color = 'rgb(200,200,200)', dash = 'dot', width = 0.5)),
      list(type = 'line', layer = 'bottom', 
                  y0 = 0, y1 = 41,
                  x0 = MeanPermCost_2, x1 = MeanPermCost_2,
                  line = list(color = 'rgb(50,50,50)', dash = 'dot', width = 0.5))
      ),
    xaxis = list(title = "Permanent Staff<br>Unit Labour Cost (USD)",
                 range = c(1,6.5), showline = TRUE, showgrid = FALSE),
    yaxis = list(title = "Farm",showticklabels = FALSE,
                 range = c(0,41), showline = TRUE, showgrid = FALSE)
  )

SingleTotalLabDumbbellPlot_SEAS <- plot_ly(data = RawData, type = 'scatter', mode = 'markers') %>%
  add_segments(x = ~SeasLabCOST_2, xend = ~SeasLabCOST_1, y = ~as.numeric(X), yend = ~as.numeric(X), 
               showlegend = FALSE,
               name = 'Seasonal',
               line = list(color = SegmentLineColor)) %>%
  add_markers(x = ~SeasLabCOST_1,
              y = ~as.numeric(X),
              name = "Pre-Covid",
              showlegend = FALSE,
              marker = list(symbol = PreMarker,
                            size = MarkerSize, 
                          color = PreMarkerColor,
                          line = list(color = 'rgb(0,0,0)'),
                          sizemode = 'diameter', opacity = 0.5)) %>%
  add_markers(name = 'Post-Covid',
              showlegend = FALSE,
            x = ~SeasLabCOST_2,
            y = ~as.numeric(X),
            marker = list(symbol = PostMarker,
                          size = MarkerSize, 
                          color = PostMarkerColor,
                          line = list(color = 'rgb(0,0,0)'),
                          sizemode = 'diameter', opacity = 0.5)) %>%
  layout(
    shapes = list(
      list(type = 'line', layer = 'bottom', 
                  y0 = 0, y1 = 41,
                  x0 = MeanSeasCost_1, x1 = MeanSeasCost_1,
                  line = list(color = 'rgb(200,200,200)', dash = 'dot', width = 0.5)),
      list(type = 'line', layer = 'bottom', 
                  y0 = 0, y1 = 41,
                  x0 = MeanSeasCost_2, x1 = MeanSeasCost_2,
                  line = list(color = 'rgb(50,50,50)', dash = 'dot', width = 0.5))
      ),
    xaxis = list(title = "Seasonal Staff<br>Unit Labour Cost (USD)",
                 range = c(2,6), showline = TRUE, showgrid = FALSE),
    yaxis = list(title = "",showticklabels = FALSE,
                 range = c(0,41), showline = TRUE, showgrid = FALSE)
  )

SingleTotalLabDumbbellPlot_NONAQ <- plot_ly(data = RawData, type = 'scatter', mode = 'markers') %>%
  add_segments(x = ~NonAqLabCOST_2, xend = ~NonAqLabCOST_1, y = ~as.numeric(X), yend = ~as.numeric(X), 
               showlegend = FALSE,
               name = 'Non-technical',
               line = list(color = SegmentLineColor)) %>%
  add_markers(x = ~NonAqLabCOST_1,
              y = ~as.numeric(X),
              name = "Pre-Covid",
              showlegend = FALSE,
              marker = list(
                          symbol = PreMarker,
                          size = MarkerSize, 
                          color = PreMarkerColor,
                          line = list(color = 'rgb(0,0,0)'),
                          sizemode = 'diameter', opacity = 0.5)) %>%
  add_markers(name = 'Post-Covid',
              showlegend = FALSE,
            x = ~NonAqLabCOST_2,
            y = ~as.numeric(X),
            marker = list(symbol = PostMarker,
                          size = MarkerSize, 
                          color = PostMarkerColor,
                          line = list(color = 'rgb(0,0,0)'),
                          sizemode = 'diameter', opacity = 0.5)) %>%
  layout(
    shapes = list(
      list(type = 'line', layer = 'bottom', 
                  y0 = 0, y1 = 41,
                  x0 = MeanNonAqCost_1, x1 = MeanNonAqCost_1,
                  line = list(color = 'rgb(200,200,200)', dash = 'dot', width = 0.5)),
      list(type = 'line', layer = 'bottom', 
                  y0 = 0, y1 = 41,
                  x0 = MeanNonAqCost_2, x1 = MeanNonAqCost_2,
                  line = list(color = 'rgb(50,50,50)', dash = 'dot', width = 0.5))
      ),
    xaxis = list(title = "Non-technical Staff<br>Unit Labour Cost (USD)",
                 range = c(2,6), showline = TRUE, showgrid = FALSE),
    yaxis = list(title = "", showticklabels = FALSE,
                 range = c(0,41), showline = TRUE, showgrid = FALSE)
  )

SingleTotalLabDumbbellPlot_Subplot <- subplot(SingleTotalLabDumbbellPlot_PERM,
                                              SingleTotalLabDumbbellPlot_SEAS,
                                              SingleTotalLabDumbbellPlot_NONAQ,
                                              titleX = TRUE, titleY = TRUE, shareY = FALSE,
                                              margin = 0.05)

```

### Total Labour Costs

The total mean farm labour costs (cost per type of worker x number of workers) was **`r mean(RawData$TotalCOST_1, na.rm = TRUE)`** per day pre-covid19 compared to **`r mean(RawData$TotalCOST_2, na.rm = TRUE)`** per day post-covid19.

For each type of employment sector this is broken down into:  

Permanent staff Pre-covid Mean total cost:  **`r mean(RawData$TotalCOST_Perm_1, na.rm = TRUE)`**
Permanent staff Post-covid Mean total cost:  **`r mean(RawData$TotalCOST_Perm_2, na.rm = TRUE)`**
  
Seasonal staff Pre-covid Mean total cost:  **`r mean(RawData$TotalCOST_Seas_1, na.rm = TRUE)`**
Seasonal staff Post-covid Mean total cost: **`r mean(RawData$TotalCOST_Seas_2, na.rm = TRUE)`**
  
Non-technical staff Pre-covid Mean total cost:  **`r mean(RawData$TotalCOST_NonAq_1, na.rm = TRUE)`**
Non-technical staff Post-covid Mean total cost: **`r mean(RawData$TotalCOST_NonAq_2, na.rm = TRUE)`**

In the following diagram the difference in total staff numbers and the difference in the total cost of labour (cost of labour x number of employees) is plotted as a bubble chart. On this chart the size of the bubble reflects the total labour cost pre-COVID19 and the colour of the bubble the total labour cost post-COVID19.  Therefore a large bubble that is coloured black reflects a farm with a relatively large pre-COVID19 labour cost and a large post-COVID19 labour cost. In contrast a small bubble with a white colour has a low pre- and post-COVID 19 labour cost.  

`r QuadrantTotalLabPlot`  
  
  The plot suggests that farms with larger costs pre-COVID19 reduced their staff numbers more resulting in relatively lower post-COVID19 costs.
  
  Also have created some dumbbell plots to show the change in labour costs pre and post covid. Here as a combined plot  

`r SingleTotalLabDumbbellPlot`  

  And, here as single plots.
  
`r SingleTotalLabDumbbellPlot_Subplot` 
  
```{r Product Prices Basic Stats, include=FALSE}
##############################################
# Include some basic stats about gate prices
##############################################

MeanGate_Pang_1 <- mean(RawData$GatePRICE_Pang_1, na.rm = TRUE)
MeanGate_Pang_2 <- mean(RawData$GatePRICE_Pang_2, na.rm = TRUE)
MeanGate_Tila_1 <- mean(RawData$GatePRICE_Tila_1, na.rm = TRUE)
MeanGate_Tila_2 <- mean(RawData$GatePRICE_Tila_2, na.rm = TRUE)
MeanGate_Carp_1 <- mean(RawData$GatePRICE_Carp_1, na.rm = TRUE)
MeanGate_Carp_2 <- mean(RawData$GatePRICE_Carp_2, na.rm = TRUE)
MeanGate_OtherCat_1 <- mean(RawData$GatePRICE_OtherCat_1, na.rm = TRUE)
MeanGate_OtherCat_2 <- mean(RawData$GatePRICE_OtherCat_2, na.rm = TRUE)

StdGate_Pang_1 <- sd(RawData$GatePRICE_Pang_1, na.rm = TRUE)
StdGate_Pang_2 <- sd(RawData$GatePRICE_Pang_2, na.rm = TRUE)
StdGate_Tila_1 <- sd(RawData$GatePRICE_Tila_1, na.rm = TRUE)
StdGate_Tila_2 <- sd(RawData$GatePRICE_Tila_2, na.rm = TRUE)
StdGate_Carp_1 <- sd(RawData$GatePRICE_Carp_1, na.rm = TRUE)
StdGate_Carp_2 <- sd(RawData$GatePRICE_Carp_2, na.rm = TRUE)
StdGate_OtherCat_1 <- sd(RawData$GatePRICE_OtherCat_1, na.rm = TRUE)
StdGate_OtherCat_2 <- sd(RawData$GatePRICE_OtherCat_2, na.rm = TRUE)


```
## Effect of Covid19 on Product Prices 

### Mean gate prices

Prior to covid19 the mean gate prices for fish were:  
  
 Pangasius: **`r MeanGate_Pang_1`** +/- **`r StdGate_Pang_1`**  
 Tilapia: **`r MeanGate_Tila_1`** +/- **`r StdGate_Tila_1`**  
 Carp: **`r MeanGate_Carp_1`** +/- **`r StdGate_Carp_1`**  
 Other catfish: **`r MeanGate_OtherCat_1`** +/- **`r StdGate_OtherCat_1`**
 
After covid19 the mean gate prices for fish were:  
  
 Pangasius: **`r MeanGate_Pang_2`** +/- **`r StdGate_Pang_2`**  
 Tilapia: **`r MeanGate_Tila_2`** +/- **`r StdGate_Tila_2`**  
 Carp: **`r MeanGate_Carp_2`** +/- **`r StdGate_Carp_2`**  
 Other catfish: **`r MeanGate_OtherCat_2`** +/- **`r StdGate_OtherCat_2`**
 
The percentage increase in prices were:  
  Pangasius: **`r (MeanGate_Pang_2 - MeanGate_Pang_1) / MeanGate_Pang_2 * 100`**  
  Tilapia: **`r (MeanGate_Tila_2 - MeanGate_Tila_1) / MeanGate_Tila_2 * 100`**  
  Carp: **`r (MeanGate_Carp_2 - MeanGate_Carp_1) / MeanGate_Carp_2 * 100`**  
  Other catfish: **`r (MeanGate_OtherCat_2 - MeanGate_OtherCat_1) / MeanGate_OtherCat_2 * 100`** 
  

  
```{r Effect on GATE PRICES BASIC STATS, include=FALSE}

############################################################
# Paired t-tests
############################################################
Pang_GATE_TTest <- t.test(RawData$GatePRICE_Pang_2,
                          RawData$GatePRICE_Pang_1,
                          alternative = "less",
                          paired = TRUE,
                          na.action = na.omit,
                          var.equal = TRUE)
Tila_GATE_TTest <- t.test(RawData$GatePRICE_Tila_2,
                          RawData$GatePRICE_Tila_1,
                          alternative = "less",
                          paired = TRUE,
                          na.action = na.omit,
                          var.equal = TRUE)
Carp_GATE_TTest <- t.test(RawData$GatePRICE_Carp_2,
                          RawData$GatePRICE_Carp_1,
                          alternative = "less",
                          paired = TRUE,
                          na.action = na.omit,
                          var.equal = TRUE)
OtherCat_GATE_TTest <- t.test(RawData$GatePRICE_OtherCat_2,
                          RawData$GatePRICE_OtherCat_1,
                          alternative = "less",
                          paired = TRUE,
                          na.action = na.omit,
                          var.equal = TRUE)

#####################################################
# Farms for each finfish
#####################################################
Number_PangFARMS <- sum(!is.na(RawData$GatePRICE_Pang_1) & !is.na(RawData$GatePRICE_Pang_2)) 
Number_TilaFARMS <- sum(!is.na(RawData$GatePRICE_Tila_1) & !is.na(RawData$GatePRICE_Tila_2)) 
Number_CarpFARMS <- sum(!is.na(RawData$GatePRICE_Carp_1) & !is.na(RawData$GatePRICE_Carp_2)) 
Number_OtherCatFARMS <- sum(!is.na(RawData$GatePRICE_OtherCat_1) & !is.na(RawData$GatePRICE_OtherCat_2)) 

```
### Number of farms

There were **`r Number_PangFARMS`** pangasius farms with pre & post-COVID gate prices, **`r Number_TilaFARMS`** tilapia farms with pre & post-COVID gate prices, **`r Number_CarpFARMS`** carp farms with pre & post-COVID gate prices and **`r Number_OtherCatFARMS`** other catfish farms with pre & post-COVID gate prices.


### Significance of changes to gate prices

To test if the decrease was statistically significant a pairwise Student's T-test was performed on the data with a single-tail (decrease).

For pangasius gate prices:  
        P-value: **`r Pang_GATE_TTest$p.value`**  
        Test Statistic: **`r Pang_GATE_TTest$statistic`**  
        Degrees of Freedom: **`r Pang_GATE_TTest$parameter`**  
        Estimated difference in means: **`r Pang_GATE_TTest$estimate`**  

For tilapia gate prices:  
        P-value: **`r Tila_GATE_TTest$p.value`**  
        Test Statistic: **`r Tila_GATE_TTest$statistic`**  
        Degrees of Freedom: **`r Tila_GATE_TTest$parameter`**  
        Estimated difference in means: **`r Tila_GATE_TTest$estimate`**  
        
For carp gate prices:  
        P-value: **`r Carp_GATE_TTest$p.value`**  
        Test Statistic: **`r Carp_GATE_TTest$statistic`**  
        Degrees of Freedom: **`r Carp_GATE_TTest$parameter`**  
        Estimated difference in means: **`r Carp_GATE_TTest$estimate`**  
        
For other catfish gate prices:  
        P-value: **`r OtherCat_GATE_TTest$p.value`**  
        Test Statistic: **`r OtherCat_GATE_TTest$statistic`**  
        Degrees of Freedom: **`r OtherCat_GATE_TTest$parameter`**  
        Estimated difference in means: **`r OtherCat_GATE_TTest$estimate`**  
        
```{r Scatter plots of Gate prices, include=FALSE}
###############################################################
# Include all prices on a single scatter plot
#
# and separately in a subplot x4
###############################################################
library(plotly)
GATEPriceScatter <- plot_ly(data = RawData, type = 'scatter', mode = 'markers') %>%
  add_lines(x = c(50,350),
            y = c(50,350),
            fill = 'tozeroy', fillcolor = 'rgba(240,240,240, 0.5)',
            showlegend = FALSE,
            line = list(color = 'rgb(200,200,200)', dash = 'dot')) %>%
  add_trace(name = '<I>Pangasius</I>',
            marker = list(symbol = 'circle', color = 'black'),
            x = ~GatePRICE_Pang_1,
            y = ~GatePRICE_Pang_2) %>%
  add_trace(name = 'Tilapia',
            marker = list(symbol = 'star', color = 'black'),
    x = ~GatePRICE_Tila_1,
            y = ~GatePRICE_Tila_2) %>%
  add_trace(name = 'Carp',
            marker = list(symbol = 'diamond', color = 'black'),
    x = ~GatePRICE_Carp_1,
            y = ~GatePRICE_Carp_2) %>%
  add_trace(name = 'Other catfish',
            marker = list(symbol = 'cross', color = 'black'),
    x = ~GatePRICE_OtherCat_1,
            y = ~GatePRICE_OtherCat_2) %>%
  layout(
    xaxis = list(title = "Gate Price Pre-COVID19",
      range = c(50,350), showline = TRUE),
    yaxis = list(title = "Gate Price Post-COVID19",
      range = c(50,350), showline = TRUE)
  )

# Subplot
GATEPriceScatter_Pang <- plot_ly(data = RawData, type = 'scatter', mode = 'markers') %>%
  add_lines(x = c(50,350),
            y = c(50,350),
            fill = 'tozeroy', fillcolor = 'rgba(240,240,240, 0.5)',
            showlegend = FALSE,
            line = list(color = 'rgb(200,200,200)', dash = 'dot')) %>%
  add_trace(name = '<I>Pangasius</I>',
            marker = list(symbol = 'circle', color = 'black'),
            x = ~GatePRICE_Pang_1,
            y = ~GatePRICE_Pang_2) %>%
  layout(
    xaxis = list(title = "",
      range = c(70,100), showline = TRUE),
    yaxis = list(title = "Gate Price Post-COVID19 (BDT)",
      range = c(70,100), showline = TRUE)
  )

GATEPriceScatter_Tila <- plot_ly(data = RawData, type = 'scatter', mode = 'markers') %>%
  add_lines(x = c(50,350),
            y = c(50,350),
            fill = 'tozeroy', fillcolor = 'rgba(240,240,240, 0.5)',
            showlegend = FALSE,
            line = list(color = 'rgb(200,200,200)', dash = 'dot')) %>%
  add_trace(name = 'Tilapia',
            marker = list(symbol = 'star', color = 'black'),
    x = ~GatePRICE_Tila_1,
            y = ~GatePRICE_Tila_2) %>%
  layout(
    xaxis = list(title = "",
      range = c(60,100), showline = TRUE),
    yaxis = list(title = "",
      range = c(60,100), showline = TRUE)
  )

GATEPriceScatter_Carp <- plot_ly(data = RawData, type = 'scatter', mode = 'markers') %>%
  add_lines(x = c(50,350),
            y = c(50,350),
            fill = 'tozeroy', fillcolor = 'rgba(240,240,240, 0.5)',
            showlegend = FALSE,
            line = list(color = 'rgb(200,200,200)', dash = 'dot')) %>%
  add_trace(name = 'Carp',
            marker = list(symbol = 'diamond', color = 'black'),
    x = ~GatePRICE_Carp_1,
            y = ~GatePRICE_Carp_2) %>%
  layout(
    xaxis = list(title = "",
      range = c(80,130), showline = TRUE),
    yaxis = list(title = "",
      range = c(80,130), showline = TRUE)
  )

GATEPriceScatter_Other <- plot_ly(data = RawData, type = 'scatter', mode = 'markers') %>%
  add_lines(x = c(50,350),
            y = c(50,350),
            fill = 'tozeroy', fillcolor = 'rgba(240,240,240, 0.5)',
            showlegend = FALSE,
            line = list(color = 'rgb(200,200,200)', dash = 'dot')) %>%
  add_trace(name = 'Other catfish',
            marker = list(symbol = 'cross', color = 'black'),
    x = ~GatePRICE_OtherCat_1,
            y = ~GatePRICE_OtherCat_2) %>%
  layout(
    xaxis = list(title = "Gate Price Pre-COVID19 (BDT)",
      range = c(210,330), showline = TRUE),
    yaxis = list(title = "",
      range = c(210,330), showline = TRUE)
  )

SubplotGATEPRICES <- subplot(GATEPriceScatter_Pang,
                             GATEPriceScatter_Tila,
                             GATEPriceScatter_Carp,
                             GATEPriceScatter_Other,
                             titleX = TRUE,
                             titleY = TRUE,
                             shareX = FALSE,
                             shareY = FALSE,
                             nrows = 2) %>%
  layout(legend = list(x = 0.05, y = 0.95, borderwidth = 0.5))

```

### Scatter Plots 

The gate prices were plotted as scatter plots of pre- and post covid 19 prices.  This shows whether there is price deflation or inflation due to the pandemic.  Will look at putting together with the consumer prices.

`r SubplotGATEPRICES`

```{r Consumer Prices of Finfish, include=FALSE}
##############################################
# Include some basic stats about gate prices
##############################################
library(data.table)
ConsumerData <- setDT(ConsumerData)

MeanMarket_Pang_1 <- mean(ConsumerData$Market_Pang_1, na.rm = TRUE)
MeanMarket_Pang_2 <- mean(ConsumerData$Market_Pang_2, na.rm = TRUE)
MeanMarket_Tila_1 <- mean(ConsumerData$Market_Tila_1, na.rm = TRUE)
MeanMarket_Tila_2 <- mean(ConsumerData$Market_Tila_2, na.rm = TRUE)
MeanMarket_Carp_1 <- mean(ConsumerData$Market_Carp_1, na.rm = TRUE)
MeanMarket_Carp_2 <- mean(ConsumerData$Market_Carp_2, na.rm = TRUE)
MeanMarket_OtherCat_1 <- mean(ConsumerData$Market_OtherCat_1, na.rm = TRUE)
MeanMarket_OtherCat_2 <- mean(ConsumerData$Market_OtherCat_2, na.rm = TRUE)

StdMarket_Pang_1 <- sd(ConsumerData$Market_Pang_1, na.rm = TRUE)
StdMarket_Pang_2 <- sd(ConsumerData$Market_Pang_2, na.rm = TRUE)
StdMarket_Tila_1 <- sd(ConsumerData$Market_Tila_1, na.rm = TRUE)
StdMarket_Tila_2 <- sd(ConsumerData$Market_Tila_2, na.rm = TRUE)
StdMarket_Carp_1 <- sd(ConsumerData$Market_Carp_1, na.rm = TRUE)
StdMarket_Carp_2 <- sd(ConsumerData$Market_Carp_2, na.rm = TRUE)
StdMarket_OtherCat_1 <- sd(ConsumerData$Market_OtherCat_1, na.rm = TRUE)
StdMarket_OtherCat_2 <- sd(ConsumerData$Market_OtherCat_2, na.rm = TRUE)

MeanMiddleMan_Pang_1 <- mean(ConsumerData[Category == "Aratdar", ]$Market_Pang_1, na.rm = TRUE)
MeanMiddleMan_Pang_2 <- mean(ConsumerData[Category == "Aratdar", ]$Market_Pang_2, na.rm = TRUE)
MeanMiddleMan_Tila_1 <- mean(ConsumerData[Category == "Aratdar", ]$Market_Tila_1, na.rm = TRUE)
MeanMiddleMan_Tila_2 <- mean(ConsumerData[Category == "Aratdar", ]$Market_Tila_2, na.rm = TRUE)
MeanMiddleMan_Carp_1 <- mean(ConsumerData[Category == "Aratdar", ]$Market_Carp_1, na.rm = TRUE)
MeanMiddleMan_Carp_2 <- mean(ConsumerData[Category == "Aratdar", ]$Market_Carp_2, na.rm = TRUE)
MeanMiddleMan_OtherCat_1 <- mean(ConsumerData[Category == "Aratdar", ]$Market_OtherCat_1, na.rm = TRUE)
MeanMiddleMan_OtherCat_2 <- mean(ConsumerData[Category == "Aratdar", ]$Market_OtherCat_2, na.rm = TRUE)

StdMiddleMan_Pang_1 <- sd(ConsumerData[Category == "Aratdar", ]$Market_Pang_1, na.rm = TRUE)
StdMiddleMan_Pang_2 <- sd(ConsumerData[Category == "Aratdar", ]$Market_Pang_2, na.rm = TRUE)
StdMiddleMan_Tila_1 <- sd(ConsumerData[Category == "Aratdar", ]$Market_Tila_1, na.rm = TRUE)
StdMiddleMan_Tila_2 <- sd(ConsumerData[Category == "Aratdar", ]$Market_Tila_2, na.rm = TRUE)
StdMiddleMan_Carp_1 <- sd(ConsumerData[Category == "Aratdar", ]$Market_Carp_1, na.rm = TRUE)
StdMiddleMan_Carp_2 <- sd(ConsumerData[Category == "Aratdar", ]$Market_Carp_2, na.rm = TRUE)
StdMiddleMan_OtherCat_1 <- sd(ConsumerData[Category == "Aratdar", ]$Market_OtherCat_1, na.rm = TRUE)
StdMiddleMan_OtherCat_2 <- sd(ConsumerData[Category == "Aratdar", ]$Market_OtherCat_2, na.rm = TRUE)


MeanRetailer_Pang_1 <- mean(ConsumerData[Category == "Paikar", ]$Market_Pang_1, na.rm = TRUE)
MeanRetailer_Pang_2 <- mean(ConsumerData[Category == "Paikar", ]$Market_Pang_2, na.rm = TRUE)
MeanRetailer_Tila_1 <- mean(ConsumerData[Category == "Paikar", ]$Market_Tila_1, na.rm = TRUE)
MeanRetailer_Tila_2 <- mean(ConsumerData[Category == "Paikar", ]$Market_Tila_2, na.rm = TRUE)
MeanRetailer_Carp_1 <- mean(ConsumerData[Category == "Paikar", ]$Market_Carp_1, na.rm = TRUE)
MeanRetailer_Carp_2 <- mean(ConsumerData[Category == "Paikar", ]$Market_Carp_2, na.rm = TRUE)
MeanRetailer_OtherCat_1 <- mean(ConsumerData[Category == "Paikar", ]$Market_OtherCat_1, na.rm = TRUE)
MeanRetailer_OtherCat_2 <- mean(ConsumerData[Category == "Paikar", ]$Market_OtherCat_2, na.rm = TRUE)

StdRetailer_Pang_1 <- sd(ConsumerData[Category == "Paikar", ]$Market_Pang_1, na.rm = TRUE)
StdRetailer_Pang_2 <- sd(ConsumerData[Category == "Paikar", ]$Market_Pang_2, na.rm = TRUE)
StdRetailer_Tila_1 <- sd(ConsumerData[Category == "Paikar", ]$Market_Tila_1, na.rm = TRUE)
StdRetailer_Tila_2 <- sd(ConsumerData[Category == "Paikar", ]$Market_Tila_2, na.rm = TRUE)
StdRetailer_Carp_1 <- sd(ConsumerData[Category == "Paikar", ]$Market_Carp_1, na.rm = TRUE)
StdRetailer_Carp_2 <- sd(ConsumerData[Category == "Paikar", ]$Market_Carp_2, na.rm = TRUE)
StdRetailer_OtherCat_1 <- sd(ConsumerData[Category == "Paikar", ]$Market_OtherCat_1, na.rm = TRUE)
StdRetailer_OtherCat_2 <- sd(ConsumerData[Category == "Paikar", ]$Market_OtherCat_2, na.rm = TRUE)

MeanConsumer_Pang_1 <- mean(ConsumerData[Category == "Consumer", ]$Market_Pang_1, na.rm = TRUE)
MeanConsumer_Pang_2 <- mean(ConsumerData[Category == "Consumer", ]$Market_Pang_2, na.rm = TRUE)
MeanConsumer_Tila_1 <- mean(ConsumerData[Category == "Consumer", ]$Market_Tila_1, na.rm = TRUE)
MeanConsumer_Tila_2 <- mean(ConsumerData[Category == "Consumer", ]$Market_Tila_2, na.rm = TRUE)
MeanConsumer_Carp_1 <- mean(ConsumerData[Category == "Consumer", ]$Market_Carp_1, na.rm = TRUE)
MeanConsumer_Carp_2 <- mean(ConsumerData[Category == "Consumer", ]$Market_Carp_2, na.rm = TRUE)
MeanConsumer_OtherCat_1 <- mean(ConsumerData[Category == "Consumer", ]$Market_OtherCat_1, na.rm = TRUE)
MeanConsumer_OtherCat_2 <- mean(ConsumerData[Category == "Consumer", ]$Market_OtherCat_2, na.rm = TRUE)

StdConsumer_Pang_1 <- sd(ConsumerData[Category == "Consumer", ]$Market_Pang_1, na.rm = TRUE)
StdConsumer_Pang_2 <- sd(ConsumerData[Category == "Consumer", ]$Market_Pang_2, na.rm = TRUE)
StdConsumer_Tila_1 <- sd(ConsumerData[Category == "Consumer", ]$Market_Tila_1, na.rm = TRUE)
StdConsumer_Tila_2 <- sd(ConsumerData[Category == "Consumer", ]$Market_Tila_2, na.rm = TRUE)
StdConsumer_Carp_1 <- sd(ConsumerData[Category == "Consumer", ]$Market_Carp_1, na.rm = TRUE)
StdConsumer_Carp_2 <- sd(ConsumerData[Category == "Consumer", ]$Market_Carp_2, na.rm = TRUE)
StdConsumer_OtherCat_1 <- sd(ConsumerData[Category == "Consumer", ]$Market_OtherCat_1, na.rm = TRUE)
StdConsumer_OtherCat_2 <- sd(ConsumerData[Category == "Consumer", ]$Market_OtherCat_2, na.rm = TRUE)
```

### Mean consumer prices

Prior to covid19 the mean consumer prices for fish were:  
  
 Pangasius: **`r MeanMarket_Pang_1`** +/- **`r StdMarket_Pang_1`**  
 Tilapia: **`r MeanMarket_Tila_1`** +/- **`r StdMarket_Tila_1`**  
 Carp: **`r MeanMarket_Carp_1`** +/- **`r StdMarket_Carp_1`**  
 Other catfish: **`r MeanMarket_OtherCat_1`** +/- **`r StdMarket_OtherCat_1`**
 
After covid19 the mean Market prices for fish were:  
  
 Pangasius: **`r MeanMarket_Pang_2`** +/- **`r StdMarket_Pang_2`**  
 Tilapia: **`r MeanMarket_Tila_2`** +/- **`r StdMarket_Tila_2`**  
 Carp: **`r MeanMarket_Carp_2`** +/- **`r StdMarket_Carp_2`**  
 Other catfish: **`r MeanMarket_OtherCat_2`** +/- **`r StdMarket_OtherCat_2`**
 
The percentage increase in prices were:  
  Pangasius: **`r (MeanMarket_Pang_2 - MeanMarket_Pang_1) / MeanMarket_Pang_2 * 100`**  
  Tilapia: **`r (MeanMarket_Tila_2 - MeanMarket_Tila_1) / MeanMarket_Tila_2 * 100`**  
  Carp: **`r (MeanMarket_Carp_2 - MeanMarket_Carp_1) / MeanMarket_Carp_2 * 100`**  
  Other catfish: **`r (MeanMarket_OtherCat_2 - MeanMarket_OtherCat_1) / MeanMarket_OtherCat_2 * 100`** 
  
The consumer data was based on **`r ConsumerData[Category == "Aratdar", .N, Category]$N`** Aratdar or Middlemen, **`r ConsumerData[Category == "Paikar", .N, Category]$N`** Paiker, or retailers, and **`r ConsumerData[Category == "Consumer", .N, Category]$N`** consumers. 
  
For each category of merchandiser the values are:

#### Middlemen

Prior to covid19 the mean consumer prices for fish were:  
  
 Pangasius: **`r MeanMiddleMan_Pang_1`** +/- **`r StdMiddleMan_Pang_1`**  
 Tilapia: **`r MeanMiddleMan_Tila_1`** +/- **`r StdMiddleMan_Tila_1`**  
 Carp: **`r MeanMiddleMan_Carp_1`** +/- **`r StdMiddleMan_Carp_1`**  
 Other catfish: **`r MeanMiddleMan_OtherCat_1`** +/- **`r StdMiddleMan_OtherCat_1`**
 
After covid19 the mean MiddleMan prices for fish were:  
  
 Pangasius: **`r MeanMiddleMan_Pang_2`** +/- **`r StdMiddleMan_Pang_2`**  
 Tilapia: **`r MeanMiddleMan_Tila_2`** +/- **`r StdMiddleMan_Tila_2`**  
 Carp: **`r MeanMiddleMan_Carp_2`** +/- **`r StdMiddleMan_Carp_2`**  
 Other catfish: **`r MeanMiddleMan_OtherCat_2`** +/- **`r StdMiddleMan_OtherCat_2`**
 
The percentage increase in prices were:  
  Pangasius: **`r (MeanMiddleMan_Pang_2 - MeanMiddleMan_Pang_1) / MeanMiddleMan_Pang_2 * 100`**  
  Tilapia: **`r (MeanMiddleMan_Tila_2 - MeanMiddleMan_Tila_1) / MeanMiddleMan_Tila_2 * 100`**  
  Carp: **`r (MeanMiddleMan_Carp_2 - MeanMiddleMan_Carp_1) / MeanMiddleMan_Carp_2 * 100`**  
  Other catfish: **`r (MeanMiddleMan_OtherCat_2 - MeanMiddleMan_OtherCat_1) / MeanMiddleMan_OtherCat_2 * 100`** 
  
#### Retailer

Prior to covid19 the mean consumer prices for fish were:  
  
 Pangasius: **`r MeanRetailer_Pang_1`** +/- **`r StdRetailer_Pang_1`**  
 Tilapia: **`r MeanRetailer_Tila_1`** +/- **`r StdRetailer_Tila_1`**  
 Carp: **`r MeanRetailer_Carp_1`** +/- **`r StdRetailer_Carp_1`**  
 Other catfish: **`r MeanRetailer_OtherCat_1`** +/- **`r StdRetailer_OtherCat_1`**
 
After covid19 the mean Retailer prices for fish were:  
  
 Pangasius: **`r MeanRetailer_Pang_2`** +/- **`r StdRetailer_Pang_2`**  
 Tilapia: **`r MeanRetailer_Tila_2`** +/- **`r StdRetailer_Tila_2`**  
 Carp: **`r MeanRetailer_Carp_2`** +/- **`r StdRetailer_Carp_2`**  
 Other catfish: **`r MeanRetailer_OtherCat_2`** +/- **`r StdRetailer_OtherCat_2`**
 
The percentage increase in prices were:  
  Pangasius: **`r (MeanRetailer_Pang_2 - MeanRetailer_Pang_1) / MeanRetailer_Pang_2 * 100`**  
  Tilapia: **`r (MeanRetailer_Tila_2 - MeanRetailer_Tila_1) / MeanRetailer_Tila_2 * 100`**  
  Carp: **`r (MeanRetailer_Carp_2 - MeanRetailer_Carp_1) / MeanRetailer_Carp_2 * 100`**  
  Other catfish: **`r (MeanRetailer_OtherCat_2 - MeanRetailer_OtherCat_1) / MeanRetailer_OtherCat_2 * 100`**   
  
#### Consumer

Prior to covid19 the mean consumer prices for fish were:  
  
 Pangasius: **`r MeanConsumer_Pang_1`** +/- **`r StdConsumer_Pang_1`**  
 Tilapia: **`r MeanConsumer_Tila_1`** +/- **`r StdConsumer_Tila_1`**  
 Carp: **`r MeanConsumer_Carp_1`** +/- **`r StdConsumer_Carp_1`**  
 Other catfish: **`r MeanConsumer_OtherCat_1`** +/- **`r StdConsumer_OtherCat_1`**
 
After covid19 the mean Consumer prices for fish were:  
  
 Pangasius: **`r MeanConsumer_Pang_2`** +/- **`r StdConsumer_Pang_2`**  
 Tilapia: **`r MeanConsumer_Tila_2`** +/- **`r StdConsumer_Tila_2`**  
 Carp: **`r MeanConsumer_Carp_2`** +/- **`r StdConsumer_Carp_2`**  
 Other catfish: **`r MeanConsumer_OtherCat_2`** +/- **`r StdConsumer_OtherCat_2`**
 
The percentage increase in prices were:  
  Pangasius: **`r (MeanConsumer_Pang_2 - MeanConsumer_Pang_1) / MeanConsumer_Pang_2 * 100`**  
  Tilapia: **`r (MeanConsumer_Tila_2 - MeanConsumer_Tila_1) / MeanConsumer_Tila_2 * 100`**  
  Carp: **`r (MeanConsumer_Carp_2 - MeanConsumer_Carp_1) / MeanConsumer_Carp_2 * 100`**  
  Other catfish: **`r (MeanConsumer_OtherCat_2 - MeanConsumer_OtherCat_1) / MeanConsumer_OtherCat_2 * 100`**    
  
### Product mark up

Based on the mean gate prices and the mean selling price the mark up of the product was calculated.

Mark up of the mean price of pangasius was:

  Pre-covid: **`r (MeanMarket_Pang_1 - MeanGate_Pang_1)/MeanGate_Pang_1 * 100`** %  
  Post-covid: **`r (MeanMarket_Pang_2 - MeanGate_Pang_2)/MeanGate_Pang_2 * 100`** %  
  Diff: **`r ((MeanMarket_Pang_2 - MeanGate_Pang_2)/MeanGate_Pang_2 * 100) - ((MeanMarket_Pang_1 - MeanGate_Pang_1)/MeanGate_Pang_1 * 100)`** %  
Mark up of the mean price of Tilapia was:

  Pre-covid: **`r (MeanMarket_Tila_1 - MeanGate_Tila_1)/MeanGate_Tila_1 * 100`** %  
  Post-covid: **`r (MeanMarket_Tila_2 - MeanGate_Tila_2)/MeanGate_Tila_2 * 100`** %  
  Diff: **`r ((MeanMarket_Tila_2 - MeanGate_Tila_2)/MeanGate_Tila_2 * 100) - ((MeanMarket_Tila_1 - MeanGate_Tila_1)/MeanGate_Tila_1 * 100)`** %  
   
Mark up of the mean price of carp was:

  Pre-covid: **`r (MeanMarket_Carp_1 - MeanGate_Carp_1)/MeanGate_Carp_1 * 100`** %  
  Post-covid: **`r (MeanMarket_Carp_2 - MeanGate_Carp_2)/MeanGate_Carp_2 * 100`** %  
  Diff: **`r ((MeanMarket_Carp_2 - MeanGate_Carp_2)/MeanGate_Carp_2 * 100) - ((MeanMarket_Carp_1 - MeanGate_Carp_1)/MeanGate_Carp_1 * 100)`** % 
  
Mark up of the mean price of other catfish was:

  Pre-covid: **`r (MeanMarket_OtherCat_1 - MeanGate_OtherCat_1)/MeanGate_OtherCat_1 * 100`** %  
  Post-covid: **`r (MeanMarket_OtherCat_2 - MeanGate_OtherCat_2)/MeanGate_OtherCat_2 * 100`** %  
   Diff: **`r ((MeanMarket_OtherCat_2 - MeanGate_OtherCat_2)/MeanGate_OtherCat_2 * 100) - ((MeanMarket_OtherCat_1 - MeanGate_OtherCat_1)/MeanGate_OtherCat_1 * 100)`** %  
   
 And, for each of the category of sellers.  

####  For Middlemen:
  
Mark up of the mean price of pangasius was:

  Pre-covid: **`r (MeanMiddleMan_Pang_1 - MeanGate_Pang_1)/MeanGate_Pang_1 * 100`** %  
  Post-covid: **`r (MeanMiddleMan_Pang_2 - MeanGate_Pang_2)/MeanGate_Pang_2 * 100`** %  
  Diff: **`r ((MeanMiddleMan_Pang_2 - MeanGate_Pang_2)/MeanGate_Pang_2 * 100) - ((MeanMiddleMan_Pang_1 - MeanGate_Pang_1)/MeanGate_Pang_1 * 100)`** %  
Mark up of the mean price of Tilapia was:

  Pre-covid: **`r (MeanMiddleMan_Tila_1 - MeanGate_Tila_1)/MeanGate_Tila_1 * 100`** %  
  Post-covid: **`r (MeanMiddleMan_Tila_2 - MeanGate_Tila_2)/MeanGate_Tila_2 * 100`** %  
  Diff: **`r ((MeanMiddleMan_Tila_2 - MeanGate_Tila_2)/MeanGate_Tila_2 * 100) - ((MeanMiddleMan_Tila_1 - MeanGate_Tila_1)/MeanGate_Tila_1 * 100)`** %  
   
Mark up of the mean price of carp was:

  Pre-covid: **`r (MeanMiddleMan_Carp_1 - MeanGate_Carp_1)/MeanGate_Carp_1 * 100`** %  
  Post-covid: **`r (MeanMiddleMan_Carp_2 - MeanGate_Carp_2)/MeanGate_Carp_2 * 100`** %  
  Diff: **`r ((MeanMiddleMan_Carp_2 - MeanGate_Carp_2)/MeanGate_Carp_2 * 100) - ((MeanMiddleMan_Carp_1 - MeanGate_Carp_1)/MeanGate_Carp_1 * 100)`** % 
  
Mark up of the mean price of other catfish was:

  Pre-covid: **`r (MeanMiddleMan_OtherCat_1 - MeanGate_OtherCat_1)/MeanGate_OtherCat_1 * 100`** %  
  Post-covid: **`r (MeanMiddleMan_OtherCat_2 - MeanGate_OtherCat_2)/MeanGate_OtherCat_2 * 100`** %  
   Diff: **`r ((MeanMiddleMan_OtherCat_2 - MeanGate_OtherCat_2)/MeanGate_OtherCat_2 * 100) - ((MeanMiddleMan_OtherCat_1 - MeanGate_OtherCat_1)/MeanGate_OtherCat_1 * 100)`** %  
   
####  For Retailers:
  
Mark up of the mean price of pangasius was:

  Pre-covid: **`r (MeanRetailer_Pang_1 - MeanGate_Pang_1)/MeanGate_Pang_1 * 100`** %  
  Post-covid: **`r (MeanRetailer_Pang_2 - MeanGate_Pang_2)/MeanGate_Pang_2 * 100`** %  
  Diff: **`r ((MeanRetailer_Pang_2 - MeanGate_Pang_2)/MeanGate_Pang_2 * 100) - ((MeanRetailer_Pang_1 - MeanGate_Pang_1)/MeanGate_Pang_1 * 100)`** %  
Mark up of the mean price of Tilapia was:

  Pre-covid: **`r (MeanRetailer_Tila_1 - MeanGate_Tila_1)/MeanGate_Tila_1 * 100`** %  
  Post-covid: **`r (MeanRetailer_Tila_2 - MeanGate_Tila_2)/MeanGate_Tila_2 * 100`** %  
  Diff: **`r ((MeanRetailer_Tila_2 - MeanGate_Tila_2)/MeanGate_Tila_2 * 100) - ((MeanRetailer_Tila_1 - MeanGate_Tila_1)/MeanGate_Tila_1 * 100)`** %  
   
Mark up of the mean price of carp was:

  Pre-covid: **`r (MeanRetailer_Carp_1 - MeanGate_Carp_1)/MeanGate_Carp_1 * 100`** %  
  Post-covid: **`r (MeanRetailer_Carp_2 - MeanGate_Carp_2)/MeanGate_Carp_2 * 100`** %  
  Diff: **`r ((MeanRetailer_Carp_2 - MeanGate_Carp_2)/MeanGate_Carp_2 * 100) - ((MeanRetailer_Carp_1 - MeanGate_Carp_1)/MeanGate_Carp_1 * 100)`** % 
  
Mark up of the mean price of other catfish was:

  Pre-covid: **`r (MeanRetailer_OtherCat_1 - MeanGate_OtherCat_1)/MeanGate_OtherCat_1 * 100`** %  
  Post-covid: **`r (MeanRetailer_OtherCat_2 - MeanGate_OtherCat_2)/MeanGate_OtherCat_2 * 100`** %  
   Diff: **`r ((MeanRetailer_OtherCat_2 - MeanGate_OtherCat_2)/MeanGate_OtherCat_2 * 100) - ((MeanRetailer_OtherCat_1 - MeanGate_OtherCat_1)/MeanGate_OtherCat_1 * 100)`** %  
   
####  For Consumers:
  
Mark up of the mean price of pangasius was:

  Pre-covid: **`r (MeanConsumer_Pang_1 - MeanGate_Pang_1)/MeanGate_Pang_1 * 100`** %  
  Post-covid: **`r (MeanConsumer_Pang_2 - MeanGate_Pang_2)/MeanGate_Pang_2 * 100`** %  
  Diff: **`r ((MeanConsumer_Pang_2 - MeanGate_Pang_2)/MeanGate_Pang_2 * 100) - ((MeanConsumer_Pang_1 - MeanGate_Pang_1)/MeanGate_Pang_1 * 100)`** %  
Mark up of the mean price of Tilapia was:

  Pre-covid: **`r (MeanConsumer_Tila_1 - MeanGate_Tila_1)/MeanGate_Tila_1 * 100`** %  
  Post-covid: **`r (MeanConsumer_Tila_2 - MeanGate_Tila_2)/MeanGate_Tila_2 * 100`** %  
  Diff: **`r ((MeanConsumer_Tila_2 - MeanGate_Tila_2)/MeanGate_Tila_2 * 100) - ((MeanConsumer_Tila_1 - MeanGate_Tila_1)/MeanGate_Tila_1 * 100)`** %  
   
Mark up of the mean price of carp was:

  Pre-covid: **`r (MeanConsumer_Carp_1 - MeanGate_Carp_1)/MeanGate_Carp_1 * 100`** %  
  Post-covid: **`r (MeanConsumer_Carp_2 - MeanGate_Carp_2)/MeanGate_Carp_2 * 100`** %  
  Diff: **`r ((MeanConsumer_Carp_2 - MeanGate_Carp_2)/MeanGate_Carp_2 * 100) - ((MeanConsumer_Carp_1 - MeanGate_Carp_1)/MeanGate_Carp_1 * 100)`** % 
  
Mark up of the mean price of other catfish was:

  Pre-covid: **`r (MeanConsumer_OtherCat_1 - MeanGate_OtherCat_1)/MeanGate_OtherCat_1 * 100`** %  
  Post-covid: **`r (MeanConsumer_OtherCat_2 - MeanGate_OtherCat_2)/MeanGate_OtherCat_2 * 100`** %  
   Diff: **`r ((MeanConsumer_OtherCat_2 - MeanGate_OtherCat_2)/MeanGate_OtherCat_2 * 100) - ((MeanConsumer_OtherCat_1 - MeanGate_OtherCat_1)/MeanGate_OtherCat_1 * 100)`** %  
   
```{r Statistical analysis of significance of changes, include=FALSE}
############################################################
# Paired t-tests - all selling prices
############################################################
Pang_MARKET_TTest <- t.test(ConsumerData$Market_Pang_2,
                          ConsumerData$Market_Pang_1,
                          alternative = "two.sided",
                          paired = TRUE,
                          na.action = na.omit,
                          var.equal = TRUE)
Tila_MARKET_TTest <- t.test(ConsumerData$Market_Tila_2,
                          ConsumerData$Market_Tila_1,
                          alternative = "two.sided",
                          paired = TRUE,
                          na.action = na.omit,
                          var.equal = TRUE)
Carp_MARKET_TTest <- t.test(ConsumerData$Market_Carp_2,
                          ConsumerData$Market_Carp_1,
                          alternative = "two.sided",
                          paired = TRUE,
                          na.action = na.omit,
                          var.equal = TRUE)
OtherCat_MARKET_TTest <- t.test(ConsumerData$Market_OtherCat_2,
                          ConsumerData$Market_OtherCat_1,
                          alternative = "two.sided",
                          paired = TRUE,
                          na.action = na.omit,
                          var.equal = TRUE)
```

### Statistical significance in changes in product selling prices


To test if differences in prices were statistically significant a pairwise Student's T-test was performed on the data with two-tails.

For pangasius Market prices:  
        P-value: **`r Pang_MARKET_TTest$p.value`**  
        Test Statistic: **`r Pang_MARKET_TTest$statistic`**  
        Degrees of Freedom: **`r Pang_MARKET_TTest$parameter`**  
        Estimated difference in means: **`r Pang_MARKET_TTest$estimate`**  

For tilapia Market prices:  
        P-value: **`r Tila_MARKET_TTest$p.value`**  
        Test Statistic: **`r Tila_MARKET_TTest$statistic`**  
        Degrees of Freedom: **`r Tila_MARKET_TTest$parameter`**  
        Estimated difference in means: **`r Tila_MARKET_TTest$estimate`**  
        
For carp Market prices:  
        P-value: **`r Carp_MARKET_TTest$p.value`**  
        Test Statistic: **`r Carp_MARKET_TTest$statistic`**  
        Degrees of Freedom: **`r Carp_MARKET_TTest$parameter`**  
        Estimated difference in means: **`r Carp_MARKET_TTest$estimate`**  
        
For other catfish Market prices:  
        P-value: **`r OtherCat_MARKET_TTest$p.value`**  
        Test Statistic: **`r OtherCat_MARKET_TTest$statistic`**  
        Degrees of Freedom: **`r OtherCat_MARKET_TTest$parameter`**  
        Estimated difference in means: **`r OtherCat_MARKET_TTest$estimate`**  

### Considering the product as a value chain

During the process, the selling of the produce may exist as a value chain going from:

Farm >> MiddleMan >> Retailer >> Consumer

In this case, we can compare the mean prices at each stage to deduce the incremental price increases made at each step before and after COVID 19.

The mean price difference for stage 1 Pangasius:
  Pre-COVID is: **`r MeanMiddleMan_Pang_1 - MeanGate_Pang_1`** or **`r (MeanMiddleMan_Pang_1 - MeanGate_Pang_1)/MeanGate_Pang_1 * 100`%** 
    
  Post-COVID is: **`r MeanMiddleMan_Pang_2 - MeanGate_Pang_2`** or **`r (MeanMiddleMan_Pang_2 - MeanGate_Pang_2)/MeanGate_Pang_2 * 100`%** 
  And the difference in markup for stage 1 is:  
  **`r (MeanMiddleMan_Pang_2 - MeanGate_Pang_2) - (MeanMiddleMan_Pang_1 - MeanGate_Pang_1)`** or a markup percentage change of: 
  **`r ((MeanMiddleMan_Pang_2 - MeanGate_Pang_2)/MeanGate_Pang_2 * 100) - ((MeanMiddleMan_Pang_1 - MeanGate_Pang_1)/MeanGate_Pang_1 * 100)`%**  
    
The mean price difference for stage 2 Pangasius:
  Pre-COVID is: **`r MeanRetailer_Pang_1 - MeanMiddleMan_Pang_1`** or **`r (MeanRetailer_Pang_1 - MeanMiddleMan_Pang_1)/MeanMiddleMan_Pang_1 * 100`%**   
  
  Post-COVID is: **`r MeanRetailer_Pang_2 - MeanMiddleMan_Pang_2`** or **`r (MeanRetailer_Pang_2 - MeanMiddleMan_Pang_2)/MeanMiddleMan_Pang_2 * 100`%**
  
  And the difference in markup for stage 2 is:  
  **`r (MeanRetailer_Pang_2 - MeanMiddleMan_Pang_2) - (MeanRetailer_Pang_1 - MeanMiddleMan_Pang_1)`** or a markup percentage change of: 
  **`r ((MeanRetailer_Pang_2 - MeanMiddleMan_Pang_2)/MeanMiddleMan_Pang_2 * 100) - ((MeanRetailer_Pang_1 - MeanMiddleMan_Pang_1)/MeanMiddleMan_Pang_1 * 100)`%**  
  
  
  The mean price difference for stage 3 Pangasius:
  Pre-COVID is: **`r MeanConsumer_Pang_1 - MeanRetailer_Pang_1`** or **`r (MeanConsumer_Pang_1 - MeanRetailer_Pang_1)/MeanRetailer_Pang_1 * 100`%**   
  
  Post-COVID is: **`r MeanConsumer_Pang_2 - MeanRetailer_Pang_2`** or **`r (MeanConsumer_Pang_2 - MeanRetailer_Pang_2)/MeanRetailer_Pang_2 * 100`%**
  
  And the difference in markup for stage 3 is:  
  **`r (MeanConsumer_Pang_2 - MeanRetailer_Pang_2) - (MeanConsumer_Pang_1 - MeanRetailer_Pang_1)`** or a markup percentage change of: 
  **`r ((MeanConsumer_Pang_2 - MeanRetailer_Pang_2)/MeanRetailer_Pang_2 * 100) - ((MeanConsumer_Pang_1 - MeanRetailer_Pang_1)/MeanRetailer_Pang_1 * 100)`%**  
  
  
  
  
  The mean price difference for stage 1 Tilapia:
  Pre-COVID is: **`r MeanMiddleMan_Tila_1 - MeanGate_Tila_1`** or **`r (MeanMiddleMan_Tila_1 - MeanGate_Tila_1)/MeanGate_Tila_1 * 100`%** 
    
  Post-COVID is: **`r MeanMiddleMan_Tila_2 - MeanGate_Tila_2`** or **`r (MeanMiddleMan_Tila_2 - MeanGate_Tila_2)/MeanGate_Tila_2 * 100`%** 
  And the difference in markup for stage 1 is:  
  **`r (MeanMiddleMan_Tila_2 - MeanGate_Tila_2) - (MeanMiddleMan_Tila_1 - MeanGate_Tila_1)`** or a markup percentage change of: 
  **`r ((MeanMiddleMan_Tila_2 - MeanGate_Tila_2)/MeanGate_Tila_2 * 100) - ((MeanMiddleMan_Tila_1 - MeanGate_Tila_1)/MeanGate_Tila_1 * 100)`%**  
    
The mean price difference for stage 2 Tilapia:
  Pre-COVID is: **`r MeanRetailer_Tila_1 - MeanMiddleMan_Tila_1`** or **`r (MeanRetailer_Tila_1 - MeanMiddleMan_Tila_1)/MeanMiddleMan_Tila_1 * 100`%**   
  
  Post-COVID is: **`r MeanRetailer_Tila_2 - MeanMiddleMan_Tila_2`** or **`r (MeanRetailer_Tila_2 - MeanMiddleMan_Tila_2)/MeanMiddleMan_Tila_2 * 100`%**
  
  And the difference in markup for stage 2 is:  
  **`r (MeanRetailer_Tila_2 - MeanMiddleMan_Tila_2) - (MeanRetailer_Tila_1 - MeanMiddleMan_Tila_1)`** or a markup percentage change of: 
  **`r ((MeanRetailer_Tila_2 - MeanMiddleMan_Tila_2)/MeanMiddleMan_Tila_2 * 100) - ((MeanRetailer_Tila_1 - MeanMiddleMan_Tila_1)/MeanMiddleMan_Tila_1 * 100)`%**  
  
  
  The mean price difference for stage 3 Tilapia:
  Pre-COVID is: **`r MeanConsumer_Tila_1 - MeanRetailer_Tila_1`** or **`r (MeanConsumer_Tila_1 - MeanRetailer_Tila_1)/MeanRetailer_Tila_1 * 100`%**   
  
  Post-COVID is: **`r MeanConsumer_Tila_2 - MeanRetailer_Tila_2`** or **`r (MeanConsumer_Tila_2 - MeanRetailer_Tila_2)/MeanRetailer_Tila_2 * 100`%**
  
  And the difference in markup for stage 3 is:  
  **`r (MeanConsumer_Tila_2 - MeanRetailer_Tila_2) - (MeanConsumer_Tila_1 - MeanRetailer_Tila_1)`** or a markup percentage change of: 
  **`r ((MeanConsumer_Tila_2 - MeanRetailer_Tila_2)/MeanRetailer_Tila_2 * 100) - ((MeanConsumer_Tila_1 - MeanRetailer_Tila_1)/MeanRetailer_Tila_1 * 100)`%** 
        
        
  
  The mean price difference for stage 1 Carp:
  Pre-COVID is: **`r MeanMiddleMan_Carp_1 - MeanGate_Carp_1`** or **`r (MeanMiddleMan_Carp_1 - MeanGate_Carp_1)/MeanGate_Carp_1 * 100`%** 
    
  Post-COVID is: **`r MeanMiddleMan_Carp_2 - MeanGate_Carp_2`** or **`r (MeanMiddleMan_Carp_2 - MeanGate_Carp_2)/MeanGate_Carp_2 * 100`%** 
  And the difference in markup for stage 1 is:  
  **`r (MeanMiddleMan_Carp_2 - MeanGate_Carp_2) - (MeanMiddleMan_Carp_1 - MeanGate_Carp_1)`** or a markup percentage change of: 
  **`r ((MeanMiddleMan_Carp_2 - MeanGate_Carp_2)/MeanGate_Carp_2 * 100) - ((MeanMiddleMan_Carp_1 - MeanGate_Carp_1)/MeanGate_Carp_1 * 100)`%**  
    
The mean price difference for stage 2 Carp:
  Pre-COVID is: **`r MeanRetailer_Carp_1 - MeanMiddleMan_Carp_1`** or **`r (MeanRetailer_Carp_1 - MeanMiddleMan_Carp_1)/MeanMiddleMan_Carp_1 * 100`%**   
  
  Post-COVID is: **`r MeanRetailer_Carp_2 - MeanMiddleMan_Carp_2`** or **`r (MeanRetailer_Carp_2 - MeanMiddleMan_Carp_2)/MeanMiddleMan_Carp_2 * 100`%**
  
  And the difference in markup for stage 2 is:  
  **`r (MeanRetailer_Carp_2 - MeanMiddleMan_Carp_2) - (MeanRetailer_Carp_1 - MeanMiddleMan_Carp_1)`** or a markup percentage change of: 
  **`r ((MeanRetailer_Carp_2 - MeanMiddleMan_Carp_2)/MeanMiddleMan_Carp_2 * 100) - ((MeanRetailer_Carp_1 - MeanMiddleMan_Carp_1)/MeanMiddleMan_Carp_1 * 100)`%**  
  
  
  The mean price difference for stage 3 Carp:
  Pre-COVID is: **`r MeanConsumer_Carp_1 - MeanRetailer_Carp_1`** or **`r (MeanConsumer_Carp_1 - MeanRetailer_Carp_1)/MeanRetailer_Carp_1 * 100`%**   
  
  Post-COVID is: **`r MeanConsumer_Carp_2 - MeanRetailer_Carp_2`** or **`r (MeanConsumer_Carp_2 - MeanRetailer_Carp_2)/MeanRetailer_Carp_2 * 100`%**
  
  And the difference in markup for stage 3 is:  
  **`r (MeanConsumer_Carp_2 - MeanRetailer_Carp_2) - (MeanConsumer_Carp_1 - MeanRetailer_Carp_1)`** or a markup percentage change of: 
  **`r ((MeanConsumer_Carp_2 - MeanRetailer_Carp_2)/MeanRetailer_Carp_2 * 100) - ((MeanConsumer_Carp_1 - MeanRetailer_Carp_1)/MeanRetailer_Carp_1 * 100)`%** 
  
    The mean price difference for stage 1 OtherCat:
  Pre-COVID is: **`r MeanMiddleMan_OtherCat_1 - MeanGate_OtherCat_1`** or **`r (MeanMiddleMan_OtherCat_1 - MeanGate_OtherCat_1)/MeanGate_OtherCat_1 * 100`%** 
    
  Post-COVID is: **`r MeanMiddleMan_OtherCat_2 - MeanGate_OtherCat_2`** or **`r (MeanMiddleMan_OtherCat_2 - MeanGate_OtherCat_2)/MeanGate_OtherCat_2 * 100`%** 
  And the difference in markup for stage 1 is:  
  **`r (MeanMiddleMan_OtherCat_2 - MeanGate_OtherCat_2) - (MeanMiddleMan_OtherCat_1 - MeanGate_OtherCat_1)`** or a markup percentage change of: 
  **`r ((MeanMiddleMan_OtherCat_2 - MeanGate_OtherCat_2)/MeanGate_OtherCat_2 * 100) - ((MeanMiddleMan_OtherCat_1 - MeanGate_OtherCat_1)/MeanGate_OtherCat_1 * 100)`%**  
    
The mean price difference for stage 2 OtherCat:
  Pre-COVID is: **`r MeanRetailer_OtherCat_1 - MeanMiddleMan_OtherCat_1`** or **`r (MeanRetailer_OtherCat_1 - MeanMiddleMan_OtherCat_1)/MeanMiddleMan_OtherCat_1 * 100`%**   
  
  Post-COVID is: **`r MeanRetailer_OtherCat_2 - MeanMiddleMan_OtherCat_2`** or **`r (MeanRetailer_OtherCat_2 - MeanMiddleMan_OtherCat_2)/MeanMiddleMan_OtherCat_2 * 100`%**
  
  And the difference in markup for stage 2 is:  
  **`r (MeanRetailer_OtherCat_2 - MeanMiddleMan_OtherCat_2) - (MeanRetailer_OtherCat_1 - MeanMiddleMan_OtherCat_1)`** or a markup percentage change of: 
  **`r ((MeanRetailer_OtherCat_2 - MeanMiddleMan_OtherCat_2)/MeanMiddleMan_OtherCat_2 * 100) - ((MeanRetailer_OtherCat_1 - MeanMiddleMan_OtherCat_1)/MeanMiddleMan_OtherCat_1 * 100)`%**  
  
  
  The mean price difference for stage 3 OtherCat:
  Pre-COVID is: **`r MeanConsumer_OtherCat_1 - MeanRetailer_OtherCat_1`** or **`r (MeanConsumer_OtherCat_1 - MeanRetailer_OtherCat_1)/MeanRetailer_OtherCat_1 * 100`%**   
  
  Post-COVID is: **`r MeanConsumer_OtherCat_2 - MeanRetailer_OtherCat_2`** or **`r (MeanConsumer_OtherCat_2 - MeanRetailer_OtherCat_2)/MeanRetailer_OtherCat_2 * 100`%**
  
  And the difference in markup for stage 3 is:  
  **`r (MeanConsumer_OtherCat_2 - MeanRetailer_OtherCat_2) - (MeanConsumer_OtherCat_1 - MeanRetailer_OtherCat_1)`** or a markup percentage change of: 
  **`r ((MeanConsumer_OtherCat_2 - MeanRetailer_OtherCat_2)/MeanRetailer_OtherCat_2 * 100) - ((MeanConsumer_OtherCat_1 - MeanRetailer_OtherCat_1)/MeanRetailer_OtherCat_1 * 100)`%** 
        